Synpotic Soil Characterization

### 2024-10-16 3-reports/manuscript_figures.Rmd
clean
add correlation matrix. new version of all analyses with only subset of variables (dropped the highly correlated variables)

```{r data_processing_functions}
make_data_subset = function(data_combined){

@@

}
```

```{r data_processing, eval=TRUE}

@@

@@

library(cluster)
library(factoextra)

@@

### Specific analytes

>>>>

```{r data_processing}
make_data_subset = function(data_combined){

@@

}

@@

## 0. Correlations

```{r gg_corr, fig.height=8, fig.width=8}

fit_correlations_function = function(dat, TITLE){
  num = 
    dat %>%       
    dplyr::select(where(is.numeric)) %>%
    drop_na()

  num_clean = 
    num %>% 
    rownames_to_column("row") %>% 
    pivot_longer(-row) %>% 
    separate(name, sep = "_", into = c("name")) %>% 
    pivot_wider() %>% 
    dplyr::select(-row)

  m = cor(num_clean)
  m = round(m, 1)

  p.mat <- ggcorrplot::cor_pmat(num_clean)
  ggcorrplot::ggcorrplot(m, type = "lower",
                         hc.order = TRUE,
                         p.mat = p.mat,
                         outline.color = "black",
                         lab = TRUE, 
                         insig = "blank",
                         #colors = c("#E46726", "white", "#6D9EC1"),
  ) +
    scale_fill_gradient2(low = "orangered2", mid = "white", high = "turquoise4", limit = c(-1, 1), midpoint = 0)+
    theme(legend.position = "none")

}

corr_all = fit_correlations_function(dat = data_wide_PCA)
#corr_wle = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "WLE"), TITLE = "WLE" )
#corr_cb = fit_correlations_function(dat = data_wide_PCA %>% filter(region == "CB"), TITLE = "CB")

corr_all
```

Highly correlated variables include:

- (+) TC & WEOC, TN
- (+) TS & Chloride, Sulfate, Na, CEC
- (+) CEC & TS, Na, K, Mg, SpConductance
- (+) Na & SpConductance, TS, K, Mg
- (+) Chloride & Sulfate, TS, CEC, SpConductance, Na
- (-) pH & WEOC, Al
- (-) Ca and Al

# VERSION 1: ALL ANALYTES

<details>
  <summary> Click to Open </summary>

@@

#library(cluster)
library(factoextra)

@@

# VERSION 2: DROPPING CORRELATED ANALYTES

```{r drop}
# drop correlated analytes

data_wide_PCA2 = data_wide_PCA %>% 
  dplyr::select(-c(TN, CEC, K, Mg, Sulfate, NO3N, pH, spConductance))
```

```{r v2-pca_input}
pca_overall = fit_pca_function(data_wide_PCA2) 
pca_overall_wle = fit_pca_function(dat = data_wide_PCA2 %>% filter(region == "WLE"))
pca_overall_cb = fit_pca_function(data_wide_PCA2 %>% filter(region == "CB"))
```

```{r v2-pca_gg, fig.width=10, fig.height=6}
## PCA plots overall
#gg_pca_overall = 
ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
         groups = as.character(pca_overall$grp$region), 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=4,stroke=1, alpha = 1,
             aes(shape = pca_overall$grp$region,
                 fill = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E"))+
  scale_fill_manual(values = c("#16879C", "#BB281E"))+
  scale_shape_manual(values = c(23, 21))+
  labs(shape="",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  guides(fill = "none")+
  NULL

```{r v2-pca_gg_regions, fig.width=10, fig.height=6}

gg_pca_wle = 
  ggbiplot(pca_overall_wle$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_wle$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 1) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_wle$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
#  xlim(-6,6)+
#  ylim(-4.5,4.5)+
  labs(shape="",
       title = "PCA: WLE",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

gg_pca_cb = 
  ggbiplot(pca_overall_cb$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall_cb$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall_cb$grp$site,
                 color = groups))+ 
  scale_color_manual(breaks = c("upland", "transition", "wte", "wc"), 
                     values = pal_transect)+
#  xlim(-5.5,7.5)+
#  ylim(-5,5)+
  labs(shape="",
       title = "PCA: CB",
       subtitle = "surface horizons")+
  theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL

```{r v2-pc_loadings}
get_pc_values = function(dat){

pc_overall = get_pc_values(data_wide_PCA2) %>% mutate(region = "combined")
pc_wle = get_pc_values(data_wide_PCA2 %>% filter(region == "WLE")) %>% mutate(region = "WLE")
pc_cb = get_pc_values(data_wide_PCA2 %>% filter(region == "CB")) %>% mutate(region = "CB")

```{r v2-pc_gg, fig.height=8, fig.width=6}

```{r v2-cluster_calc, include=FALSE}

#library(cluster)
library(factoextra)

df = data_wide_PCA2 %>%   drop_na() %>% dplyr::select(where(is.numeric))
grp = data_wide_PCA2 %>% drop_na() %>% dplyr::select(!where(is.numeric))

```{r v2-cluster_gg}
df_cl %>%
  ggplot(aes(y = cluster, x = site))+
  geom_point(aes(color = transect, shape = transect),
             size = 4, stroke = 1,
             position = position_jitterdodge(jitter.width = 0.0, jitter.height = 0.1, 
                                             dodge.width = 0.4))+
  scale_color_manual(values = pal_transect)

```{r v2-cluster_pca, fig.width=10, fig.height=6}
library(ggConvexHull)

adonis2((data_wide_PCA2 %>% dplyr::select(where(is.numeric)) %>% drop_na()) ~ 
         (region + transect + horizon + site)^2, 
       data = data_wide_PCA2 %>% drop_na)

```{r v2-scaled_gg1, fig.height = 8, fig.width = 7, eval=FALSE}
data_long = 
  data_combined_wide %>% 
  rename(Cl = Chloride,
         SO4 = Sulfate) %>% 
  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

```{r v2-scaled_gg2, fig.height = 8, fig.width = 8}
data_long_all = 
  data_combined_wide %>% 
  dplyr::select(-c(TN, CEC, K, Mg, Sulfate, NO3N, pH, spConductance, SOM, DIC, Al)) %>% 
  mutate(transect = case_when(transect == "wc" ~ "wetland", TRUE ~ transect)) %>% 
  pivot_longer(cols = -c(sample_label, region, site, transect, tree_number, horizon)) %>% 
  filter(!is.na(value))

data_scaled_NEW = 
  data_long_all %>% 
  group_by(region, site, horizon, transect, name) %>% 
  dplyr::summarise(value_scaled = median(value)) %>% 
  group_by(region, site, horizon, name) %>% 
  dplyr::mutate(value_scaled = scales::rescale(value_scaled)) %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wetland"))) %>% 
  mutate(name = factor(name, levels = c("TC", "TS", "WEOC",
                                        "Ca", "Na", "Chloride",
                                        "NH4N", "Fe", "P")))

```{r v2-scaled_gg2_split, fig.height = 10, fig.width = 14}
library(ggh4x) # for nested facets
# https://teunbrand.github.io/ggh4x/reference/facet_nested.html

data_scaled_NEW %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
#  facet_wrap( ~ name)+
  facet_nested_wrap(~ name + region, ncol = 6,
                    nest_line = element_line(colour = "grey"))+ 
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())
```

```{r v2-scaled_gg2_gwi, fig.height = 8, fig.width = 8}
data_scaled_NEW %>% 
  filter(site == "GWI") %>% 
  ggplot(aes(x = transect, y = value_scaled, group = site, color = site))+
  geom_line(linewidth = 1) + geom_point(size = 2)+
  facet_wrap( ~ name )+
  scale_color_manual(breaks = c("CRC", "PTR", "OWC",
                                "GCW", "MSM", "GWI"),
                     values = c("#d73027", "#fc8d59", "#fdbb84", 
                                "#74a9cf", "#2b8cbe", "#045a8d"))+
  scale_x_discrete(labels = c("U", "T" , "W" ))+
  labs(x = "Transect", y = "Normalized concentration")+
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank())

# Specific analytes
